### 2016-02-16 paper/paper.tex
Added architecture, seawall, acknowledgement.

\section{Introduction}
\paragraph{} Simulations of natural phenomena are a precious tool for analysis and
understanding of the processes behind them as well as the implications of
those. Especially fluid dynamics is one of the fields most benefitted by the explosion of
high performance parallel computing architectures of the last years. Tsunamis are one of
the most devastating natural disasters, with much attention drawn to them lately,
especially after the 2004 Indian Ocean tsunami and the 2011 T\={o}hoku earthquake and
tsunami, two of the largest incidents of modern history. Multiscale modelling of tsunami
generation, propagation and impact is a trending research area as respective simulations
give valuable insights into the underlying mechanisms and relations between the various
stages of an unfolding tsunami incident, while also facilitating the assessment of
potential hazard it poses upon impact on a coastline.

@@

\section{Related Work}
\paragraph{} Computational fluid dynamics is a very active area of research, as it relates
to a wide spectrum of applications in computer graphics, engineering and science. The SPH
method was developed independently by Gingold and Monaghan \cite{gingold1977375} and Lucy
\cite{lucy19771013} in 1977. Since then it has been employed in a wide variety of problems
and applications, proving itself to be a flexible and attractive method for the simulation
of complex, multicomponent/multiphase flows. Its flexibility is shown by the numerous
adaptations and customizations it has undergone to suit many diverse problem domains.

Relating to the method itself, Desbrun and Gascuel \cite{desbrun1996smoothed} used the
method to animate highly deformable bodies of various stiffness and viscosity, while
proposing important extensions like the ``spiky'' pressure smoothing kernel and discussing
implementation issues such as the surface reconstruction from density isosurfaces. Later,
M\"{u}ller et al. \cite{muller2003particle} used this work as a basis for fluid simulation
in interactive applications. Becker and Teschner \cite{becker2007weakly} developed Weakly
Compressible SPH, where the ideal gas equation of state is replaced with the much more
strict Tait equation to reduce compressibility to a user-defined upper bound, thus
avoiding an inefficient explicit Poisson equation solver. In recent years, Solenthaler and
Pajarola \cite{solenthaler2009predictive} improved WCSPH by proposing
Predictive-Corrective Incompressible SPH, where a prediction-correction scheme is used to
compute particle pressures through iterative density constraint satisfaction and
propagation through the fluid, until the final tolerance conditions are met. Implicit
Incompressible SPH was then proposed by Ihmsen et al. \cite{ihmsen2014implicit}, in which
the density is predicted using a discretized form of the pressure Poisson equation, as
obtained from the combination of a direct discretization of the continuity equation and
symmetric pressure forces.

@@

\section{Proposed framework}
\paragraph{} Our proposed simulation framework is based on SPH, a lagrangian fluid
simulation method whose core notion is the discretization of the fluid into particles,
which serve as interpolation points for the estimation of fluid properties in
space. Advantages of this method include the exact treatment of advection, the natural way
of dealing with special interface interactions, the inherent conservation of significant
quantities (mass, momentum, energy) and the self-adaptivity of computational load to the
fluid location and state in the flow domain. To efficiently ensure compressibility in
degenerate cases and undersampled boundary regions while maintaining relatively large
timesteps, SPH was enhanced with explicit solving of geometric constraints between
particles. These constraints are solved by the Bullet physics engine, within which fluid
particles are represented as rigid bodies, while also being subject to SPH forces as
computed at each timestep of the simulation.

\subsection{SPH method}
\paragraph{} Starting from the identity:
\begin{equation*}
  f(\vec{r}) = \int_Vf(\vec{x}) \delta(\vec{r} - \vec{x}) d\vec{x},
\end{equation*}
where $\delta(\vec{r})$ the Dirac delta function and $\vec{x} \in V$, one can obtain a
more general interpolation rule by substituting $\delta(\vec{r})$ with a smoothing kernel
$W(\vec{r}, h)$:
\begin{equation*}
  \label{eq:c-approx}
  f(\vec{r}) \approx \int_V f(\vec{x}) W(\vec{r}-\vec{x}, h) d\vec{x}
\end{equation*}
whose limit when $h\to0$ approaches the delta fucntion and is normalized to unity:
\begin{equation*}
  \label{eq:kernel-properties}
  \lim_{h\to0}W(\vec{r}, h) = \delta(\vec{r})\ \ \
  \text{and}\ \ \
  \int_VW(\vec{r}, h) d\vec{x} = 1
\end{equation*}
The smoothing radius $h$ serves as a cutoff radius in the smoothing process, as particles
beyond that distance have no contribution to the sum, i.e. $W(r,h) = 0$ when $r>h$. For
the discrete case, where $f$ is discretized to particles with density $\rho$ and mass $m$,
the weighting ratio $m/\rho$ can be used to construct a weighted sum interpolant for any
field $A$:
\begin{equation}
  \label{eq:d-approx}
  A(\vec{r}) = \sum_i \frac{m_i}{\rho_i} A(\vec{r}_i) W(\vec{r}-\vec{x}_i, h)
\end{equation}
which lies at the heart of SPH formulation. According to this, the gradient can be
computed by the following approximation:
\begin{equation}
  \label{eq:d-grad}
  \nabla A(\vec{r}) = \sum_i \frac{m_i}{\rho_i} A(\vec{r}_i) \nabla W(\vec{r} - \vec{x}_i, h)
\end{equation}
The obvious advantage of this is the exclusive dependence on the smoothing kernel
gradient, which can be precomputed for sensible kernel choices. However, this formula can
lead to unsymmetric pair forces, compromising the conservation of linear and angular
momentum of the system. To symmetrize these forces depending on gradients (like those
originating from pressure differences), we can use the product rule:
\begin{equation*}
  \label{eq:grad-identity}
  \nabla \left( \frac{P}{\rho} \right) =
  \frac{\nabla P}{\rho}-
  \frac{P}{\rho^2} \nabla \rho
  \hspace{10pt} \Leftrightarrow \hspace{10pt}
  \nabla P = \rho \left[ \frac{P}{\rho^2} \nabla \rho + \nabla \left( \frac{P}{\rho}
    \right) \right]
\end{equation*}
to obtain an alternative approximation of gradient
\begin{align}
  \label{eq:grad-est}
  \nabla P & = \rho \left[ \frac{P}{\rho^2} \sum_i \frac{m_i}{\rho_i} \rho \nabla
             W(\vec{r}-\vec{r}_i, h)
             +
             \sum_i \frac{m_i}{\rho_i} \frac{P_i}{\rho_i} \nabla W(\vec{r}-\vec{r}_i, h)
             \right] \nonumber \\
           & = \rho \sum_i m_i \left(\frac{P}{\rho^2} + \frac{P_i}{\rho_i} \right)
             \nabla W(\vec{r}-\vec{r}_i, h)
\end{align}
which is antisymmetric for all interacting particle pairs. Viscosity forces on the other
hand, are proportional to the laplacian of the velocity field:
\begin{equation}
  \label{eq:lapl-est}
  \nabla^2\vec{v} = \sum_i \frac{m_i}{\rho_i} (\vec{v}_i - \vec{v}) \nabla^2
  W(\vec{r}-\vec{r}_i, h)
\end{equation}
and are always antisymmetric, since they depend on velocity difference
$\vec{v}_i - \vec{v}$ between particles.  In each timestep of the simulation, the density
of all particles is first computed according to equation \ref{eq:d-grad}, as it depends
only on the relative position of those. The pressure at each particle location is then
obtained from its respective density through an equation of state. Subsequently, the
pressure and viscosity forces are computed from the particle data and integrated back into
the position and velocity of the particles.

\subsection{Implementation}
\paragraph{} Each simulation under our implementation consists of two elements, the
coastline terrain (static) and the tsunami wave (dynamic). For the simulation setup,
terrain is imported from a suitable 3D geometry definition file format, the initial
conditions of the impacting wave (position and velocity) are configured, and the desired
discretization resolution for the fluid is set. From these conditions, the terrain and
fluid are initialized, and key parameters are computed as subsequently described. Terrain
geometry is scaled by a user-supplied factor and docked to the origin of coordinates,
while fluid particles are initially placed into an Hexagonal Close-Packed lattice, to
achieve the densest possible packing and symmetry:
\begin{equation}
  \label{eq:hcp}
  [x, y, z] = \left[ 2i+[(j+k) \bmod 2], \sqrt{3}[j + \frac{1}{3} (k \bmod 2)],
    \frac{2\sqrt{6}}{3} k \right] r
\end{equation}
In this configuration, the smoothing radius $h$ is computed such that each particle has
approximately 50 neighbours (following the empirical rule established in the
literature). The timestep is determined according to the Courant-Friedrichs-Lewy
criterion:
\begin{equation}
  \label{eq:cfl}
  \delta t_{\text{CFL}} = C \frac {\delta x}{v_{\text{max}}},
\end{equation}
for values of Courant number $C\approx0.5$ with characteristic length $\delta x$ equal to
the effective particle radius and maximum velocity $v_{\text{max}}$ determined by the
maximum potential energy in the initial configuration. Following standard practice, we
used three different smoothing kernels for density, pressure gradient and velocity
laplacian computation:
\begin{equation}
  \label{eq:poly6}
  W_{\text{poly6}}(r, h) = \frac{315}{64 \pi h^9}
  \begin{cases}
    (h^2 - r^2)^3 & 0 \leq r \leq h\\
    0 & \text{otherwise}
  \end{cases}
\end{equation}
\begin{equation}
  \label{eq:spiky}
  W_{\text{spiky}}(r, h) = \frac{15}{\pi h^6}
  \begin{cases}
    (h - r)^3 & 0 \leq r \leq h\\
    0 & \text{otherwise}
  \end{cases}
\end{equation}
\begin{equation*}
  \label{eq:spiky-gradient}
  \nabla W_{\text{spiky}}(r, h) = \frac{-45}{\pi h^6} (h-r)^2
\end{equation*}
\begin{equation}
  \label{eq:viscosity}
  W_{\text{viscosity}}(r, h) = \frac{15}{2 \pi h^3}
  \begin{cases}
    - \frac{r^3}{2h^3} + \frac{r^2}{h^2} + \frac{h}{2r} - 1 & 0 \leq r \leq h\\
    0 & \text{otherwise},
  \end{cases}\\
\end{equation}
\begin{equation*}
  \label{eq:viscosity-laplacian}
  \nabla^2 W_{\text{viscosity}}(r, h) = \frac{45}{\pi h^6} (h-r)
\end{equation*}

@@

\section{Results}
\paragraph{} Multiple simulations were carried out using different models of urban
coastline, in order to gain a significant and diverse dataset of impulses exerted over the
duration of the impact. Tsunamis are vastly different from the usual wind-induced sea
waves in that they have far longer wavelength and carry much greater total energy,
appearing as a rapidly rising tide instead of a breaking waves. Accounting for these
facts, we chose to represent the tsunami wave as a water volume invading the coastline
with an initial velocity.

\begin{figure}[h!]
  \includegraphics[width=\textwidth]{figures/paraview.png}
  \caption{Example visualization of impact data in Paraview\textsuperscript{TM}. Fluid
    surface is reconstructed as isosurface of fluid density, while impulses on the
    selected region are visualized as points on the 3D terrain model and in a histogram
    grouped by their magnitude.}
  \label{fig:visualization}
\end{figure}

@@

\section{Conclusions}

>>>>

\section{Introduction}
Simulations of natural phenomena are a precious tool for analysis and understanding of the
processes behind them as well as the implications of those. Especially fluid dynamics is
one of the fields most benefitted by the explosion of high performance parallel computing
architectures of the last years. Tsunamis are one of the most devastating natural
disasters, with much attention drawn to them lately, especially after the 2004 Indian
Ocean tsunami and the 2011 T\={o}hoku earthquake and tsunami, two of the largest incidents
of modern history. Multiscale modelling of tsunami generation, propagation and impact is a
trending research area as respective simulations give valuable insights into the
underlying mechanisms and relations between the various stages of an unfolding tsunami
incident, while also facilitating the assessment of potential hazard it poses upon impact
on a coastline.

@@

\section{Related Work}
Computational fluid dynamics is a very active area of research, as it relates to a wide
spectrum of applications in computer graphics, engineering and science. The SPH method was
developed independently by Gingold and Monaghan \cite{gingold1977375} and Lucy
\cite{lucy19771013} in 1977. Since then it has been employed in a wide variety of problems
and applications, proving itself to be a flexible and attractive method for the simulation
of complex, multicomponent/multiphase flows. Its flexibility is shown by the numerous
adaptations and customizations it has undergone to suit many diverse problem domains.

Relating to the method itself, Desbrun and Gascuel \cite{desbrun1996smoothed} used it to
animate highly deformable bodies of various stiffness and viscosity, while proposing
important extensions like the ``spiky'' pressure smoothing kernel and discussing
implementation issues such as the surface reconstruction from density isosurfaces. Later,
M\"{u}ller et al. \cite{muller2003particle} used this work as a basis for fluid simulation
in interactive applications. Becker and Teschner \cite{becker2007weakly} developed Weakly
Compressible SPH, where the ideal gas equation of state is replaced with the much more
strict Tait equation to reduce compressibility to a user-defined upper bound, thus
avoiding an inefficient explicit Poisson equation solver. In recent years, Solenthaler and
Pajarola \cite{solenthaler2009predictive} improved WCSPH by proposing
Predictive-Corrective Incompressible SPH, where a prediction-correction scheme is used to
compute particle pressures through iterative density constraint satisfaction and
propagation through the fluid, until the final tolerance conditions are met. Implicit
Incompressible SPH was then proposed by Ihmsen et al. \cite{ihmsen2014implicit}, in which
the density is predicted using a discretized form of the pressure Poisson equation, as
obtained from the combination of a direct discretization of the continuity equation and
symmetric pressure forces.

@@

\section{Proposed framework}
Our proposed simulation framework is based on SPH, a lagrangian fluid simulation method
whose core notion is the discretization of the fluid into particles, which serve as
interpolation points for the estimation of fluid properties in space. Advantages of this
method include the exact treatment of advection, the natural way of dealing with special
interface interactions, the inherent conservation of significant quantities (mass,
momentum, energy) and the self-adaptivity of computational load to the fluid location and
state in the flow domain. To efficiently ensure compressibility in degenerate cases and
undersampled boundary regions while maintaining relatively large timesteps, SPH was
enhanced with explicit solving of geometric constraints between particles. These
constraints are solved by the Bullet physics engine, within which fluid particles are
represented as rigid bodies, while also being subject to SPH forces as computed at each
timestep of the simulation.

\subsection{SPH method}
Starting from the identity:
\begin{equation*}
  f(\vec{r}) = \int_Vf(\vec{x}) \delta(\vec{r} - \vec{x}) d\vec{x},
\end{equation*}
where $\delta(\vec{r})$ the Dirac delta function and $\vec{x} \in V$, one can obtain a
more general interpolation rule by substituting $\delta(\vec{r})$ with a smoothing kernel
$W(\vec{r}, h)$:
\begin{equation*}
  \label{eq:c-approx}
  f(\vec{r}) \approx \int_V f(\vec{x}) W(\vec{r}-\vec{x}, h) d\vec{x}
\end{equation*}
whose limit when $h\to0$ approaches the delta fucntion and is normalized to unity:
\begin{equation*}
  \label{eq:kernel-properties}
  \lim_{h\to0}W(\vec{r}, h) = \delta(\vec{r})\ \ \
  \text{and}\ \ \
  \int_VW(\vec{r}, h) d\vec{x} = 1
\end{equation*}
The smoothing radius $h$ serves as a cutoff radius in the smoothing process, as particles
beyond that distance have no contribution to the sum, i.e. $W(r,h) = 0$ when $r>h$. For
the discrete case, where $f$ is discretized to particles with density $\rho$ and mass $m$,
the weighting ratio $m/\rho$ can be used to construct a weighted sum interpolant for any
field $A$:
\begin{equation}
  \label{eq:d-approx}
  A(\vec{r}) = \sum_i \frac{m_i}{\rho_i} A(\vec{r}_i) W(\vec{r}-\vec{x}_i, h)
\end{equation}
which lies at the heart of SPH formulation. According to this, the gradient can be
computed by the following approximation:
\begin{equation}
  \label{eq:d-grad}
  \nabla A(\vec{r}) = \sum_i \frac{m_i}{\rho_i} A(\vec{r}_i) \nabla W(\vec{r} - \vec{x}_i, h)
\end{equation}
The obvious advantage of this is the exclusive dependence on the smoothing kernel
gradient, which can be precomputed for sensible kernel choices. However, this formula can
lead to unsymmetric pair forces, compromising the conservation of linear and angular
momentum of the system. To symmetrize these forces depending on gradients (like those
originating from pressure differences), we can use the product rule:
\begin{equation*}
  \label{eq:grad-identity}
  \nabla \left( \frac{P}{\rho} \right) =
  \frac{\nabla P}{\rho}-
  \frac{P}{\rho^2} \nabla \rho
  \hspace{10pt} \Leftrightarrow \hspace{10pt}
  \nabla P = \rho \left[ \frac{P}{\rho^2} \nabla \rho + \nabla \left( \frac{P}{\rho}
    \right) \right]
\end{equation*}
to obtain an alternative approximation of gradient
\begin{align}
  \label{eq:grad-est}
  \nabla P & = \rho \left[ \frac{P}{\rho^2} \sum_i \frac{m_i}{\rho_i} \rho \nabla
             W(\vec{r}-\vec{r}_i, h)
             +
             \sum_i \frac{m_i}{\rho_i} \frac{P_i}{\rho_i} \nabla W(\vec{r}-\vec{r}_i, h)
             \right] \nonumber \\
           & = \rho \sum_i m_i \left(\frac{P}{\rho^2} + \frac{P_i}{\rho_i} \right)
             \nabla W(\vec{r}-\vec{r}_i, h)
\end{align}
which is antisymmetric for all interacting particle pairs. Viscosity forces on the other
hand, are proportional to the laplacian of the velocity field:
\begin{equation}
  \label{eq:lapl-est}
  \nabla^2\vec{v} = \sum_i \frac{m_i}{\rho_i} (\vec{v}_i - \vec{v}) \nabla^2
  W(\vec{r}-\vec{r}_i, h)
\end{equation}
and are always antisymmetric, since they depend on velocity difference
$\vec{v}_i - \vec{v}$ between particles.  In each timestep of the simulation, the density
of all particles is first computed according to equation \ref{eq:d-grad}, as it depends
only on the relative position of those. The pressure at each particle location is then
obtained from its respective density through an equation of state. Subsequently, the
pressure and viscosity forces are computed from the particle data and integrated back into
the position and velocity of the particles.

\subsection{Implementation}
Each simulation under our implementation consists of two elements, the coastline terrain
(static) and the tsunami wave (dynamic). For the simulation setup, terrain is imported
from a suitable 3D geometry definition file format, the initial conditions of the
impacting wave (position and velocity) are configured, and the desired discretization
resolution for the fluid is set. From these conditions, the terrain and fluid are
initialized, and key parameters are computed as subsequently described. Terrain geometry
is scaled by a user-supplied factor and docked to the origin of coordinates, while fluid
particles are initially placed into an Hexagonal Close-Packed lattice, to achieve the
densest possible packing and symmetry:
\begin{equation}
  \label{eq:hcp}
  [x, y, z] = \left[ 2i+[(j+k) \bmod 2], \sqrt{3}[j + \frac{1}{3} (k \bmod 2)],
    \frac{2\sqrt{6}}{3} k \right] r
\end{equation}
In this configuration, the smoothing radius $h$ is computed such that each particle has
approximately 50 neighbours (following the empirical rule established in the
literature). The timestep is determined according to the Courant-Friedrichs-Lewy
criterion:
\begin{equation}
  \label{eq:cfl}
  \delta t_{\text{CFL}} = C \frac {\delta x}{v_{\text{max}}},
\end{equation}
for values of Courant number $C\approx0.5$ with characteristic length $\delta x$ equal to
the effective particle radius and maximum velocity $v_{\text{max}}$ determined by the
maximum potential energy in the initial configuration. Following standard practice, we
used three different smoothing kernels for density, pressure gradient and velocity
laplacian computation:
\begin{equation}
  \label{eq:poly6}
  W_{\text{poly6}}(r, h) = \frac{315}{64 \pi h^9}
  \begin{cases}
    (h^2 - r^2)^3 & 0 \leq r \leq h\\
    0 & \text{otherwise}
  \end{cases}
\end{equation}
\begin{equation}
  \label{eq:spiky}
  W_{\text{spiky}}(r, h) = \frac{15}{\pi h^6}
  \begin{cases}
    (h - r)^3 & 0 \leq r \leq h\\
    0 & \text{otherwise}
  \end{cases}
\end{equation}
\begin{equation*}
  \label{eq:spiky-gradient}
  \nabla W_{\text{spiky}}(r, h) = \frac{-45}{\pi h^6} (h-r)^2
\end{equation*}
\begin{equation}
  \label{eq:viscosity}
  W_{\text{viscosity}}(r, h) = \frac{15}{2 \pi h^3}
  \begin{cases}
    - \frac{r^3}{2h^3} + \frac{r^2}{h^2} + \frac{h}{2r} - 1 & 0 \leq r \leq h\\
    0 & \text{otherwise},
  \end{cases}\\
\end{equation}
\begin{equation*}
  \label{eq:viscosity-laplacian}
  \nabla^2 W_{\text{viscosity}}(r, h) = \frac{45}{\pi h^6} (h-r)
\end{equation*}

@@

\section{Results}
Multiple simulations were carried out using different models of urban coastline, in order
to gain a significant and diverse dataset of impulses exerted over the duration of the
impact. Tsunamis are vastly different from the usual wind-induced sea waves in that they
have far longer wavelength and carry much greater total energy, appearing as a rapidly
rising tide instead of a breaking waves. Accounting for these facts, we chose to represent
the tsunami wave as a water volume invading the coastline with an initial velocity.

\begin{figure}[h!]
  \includegraphics[width=\textwidth]{figures/paraview.png}
  \caption{Example visualization of impact data in Paraview\textsuperscript{TM}. Fluid
    surface is reconstructed as isosurface of fluid density, while impulses on the
    selected region are vi\-su\-a\-li\-zed as points on the 3D terrain model and in a
    histogram grouped by their magnitude.}
  \label{fig:visualization}
\end{figure}

@@

\section{Conclusions}
We presented the theoretical background and implementation of an enriched SPH framework,
whose main purpose was the faithful simulation and detailed recording of the forces
exerted upon the coastline during a tsunami impact. The chosen output file format lends
itself for rich visualizations, allowing for a quick overview of complex simulation data.

### 2016-02-16 report/acknowledgements.tex
Added architecture, seawall, acknowledgement.

\section*{Ευχαριστίες}

>>>>

\section*{Ευχαριστίες}
\paragraph{} Με την ολοκλήρωση της διπλωματικής αυτής εργασίας, θεωρώ σημαντικό να
ευχαριστήσω από τη θέση αυτή όλους όσους με βοήθησαν να φτάσω μέχρι εδώ. Ευχαριστώ κατ'
αρχάς τον επιβλέποντα της διπλωματικής, κύριο Κωνσταντίνο Μουστάκα. Μέσα από τη θετική του
στάση απέναντι σε οποιαδήποτε πρόταση, τις ενεργές προσπάθειές του για βοήθεια και
καθοδήγηση και την προσήλωσή του στην ουσία έχει κερδίσει την ειλικρινή προσωπική μου
εκτίμηση τον καιρό που είχα την τύχη να συνεργάζομαι μαζί του. Θα ήθελα επίσης να αναφερθώ
στη μνήμη του εκλιπόντα καθηγητή Κωνσταντίνου Ευσταθίου, ο οποίος έχοντας εμπνεύσει με την
αγάπη του προς τους φοιτητές και το αντικείμενο των σπουδών μας, αποτελούσε κατά τη γνώμη
μου μία από τις κορυφαίες προσωπικότητες του τμήματος, εξίσου ως επιστήμονας και άνθρωπος,
δυστυχώς όμως για όλους έφυγε νωρίς. Καθώς η εργασία αυτή σηματοδοτεί και το τέλος των
σπουδών μου, οφείλω να ευχαριστήσω τους ανθρώπους που με βοήθησαν. Θα ήθελα από καρδιάς να
ευχαριστήσω το φίλο και συμφοιτητή μου \eng{Dimitar Stanev} για τη φιλία και την
ανεκτίμητη και ανιδιοτελή βοήθειά του προς εμένα κατά την προετοιμασία για εργαστήρια και
εξετάσεις, όπως και τον Κωνσταντίνο Καρανικολή. Ένα ιδιαίτερο ευχαριστώ οφείλω και στην
Ειρήνη Τζιλιγκάκη για τη στήριξη, κατανόηση και την αγάπη που μου χαρίζει τα τελευταία
χρόνια. Τέλος και πάνω απ' όλους, ευγνωμονώ τους γονείς μου, που μου έδωσαν μια δεύτερη
ευκαιρία να σπουδάσω αυτό που ήθελα, καθώς πάντα ευόδωναν και υποστήριζαν τις επιλογές μου
έμπρακτα και με κάθε τρόπο.

\clearpage

\begin{center}
  \null\vfill
  \large{Αφιερωμένο στον πατέρα μου, Γεράσιμο}
  \vspace{2cm}
  \null\vfill
\end{center}

\thispagestyle{empty}
